Approximation as Differentiation: Notes

### 2025-05-26 main/category-of-families.tex
Note on reindexing functor for f, which is just the image of f in the indexed category [-,C].

The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$,
sometimes called the \emph{total category} for $T$, is the category of all set-indexed families of objects of
$T$, together with morphisms that account for both structure within each family and how families transform
under reindexing functions $f: X \to Y$. First we formalise $X$-indexed families and their categories.

@@

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
$\reindex{\eta}{f}(x) = \eta(f(x))$.
\end{definition}

@@

\begin{definition}
\label{def:Grothendieck}
\todo{Move earlier because doesn't assume $\cat{C} = \Set$.} Suppose an indexed category $T: \cat{C}^\op \to
\Cat$. The \emph{Grothendieck construction} for $T$ is the category $\Grothendieck{X}T(X)$ (also written
$\Grothendieck{}T$) which has as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an
object $\partial X$ of $T(X)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial
f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in
$T(X)$.
\end{definition}

\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:

Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
$f: X \to Y$ to the reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .

>>>>

The general form of the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
\cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, assembles all the categories $T(X)$
(for objects $X$ in $\cat{C}$) into a single category, together with morphisms that account for both internal
structure and reindexing along morphisms $f: X \to Y$ in $\cat{C}$.

\begin{definition}
\label{def:Grothendieck}
Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
\end{definition}

\noindent The $\partial X$ notation we adopt here is intended to suggest the connection between the
Grothendieck construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$
components compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g):
(Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:

Now we formalise $X$-indexed families as functor categories $\Func{X}{\cat{C}}$ and consider the specific case
of the Grothendieck construction for set-indexed categories $\Func{-}{\cat{C}}: \Set^\op \to \Cat$.

@@

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
$\reindex{\eta}{f}(x) = \eta(f(x))$. \todo{Should be clearer that $\reindex{-}{f}$ is actually
$\Func{f}{\cat{C}}$.}
\end{definition}

@@

The Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed category which sends
any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the
reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$, is called the \emph{families}
construction for $\cat{C}$.
